
\chapter{Hypothesis testing\label{chap:hypothesisTesting}}
\begin{example}
\label{exa:isi2004samplepsb9}Let $Y_{1},Y_{2},Y_{3}$ and $Y_{4}$
be a random sample from a population with probability density function
\[
f(y,\theta)=\begin{cases}
\left(\frac{1}{2\theta^{3}}\right)y^{2}\exp(-y/\theta) & \text{ if }y>0\\
0 & \text{ otherwise, }
\end{cases}
\]
 where $\theta>0$. Find the most powerful test for testing $H_{0}:\theta=\theta_{0}$
against the hypothesis $H_{1}:\theta=\theta_{1}$, where $\theta_{1}>\theta_{0}$.
Is the test uniformly most powerful for $\theta>\theta_{0}$ ?\hl{TODO}
\end{example}

\begin{example}
\label{exa:isi2004samplepsb10}On a particular day let $X_{1},X_{2}$
and $X_{3}$ be the number of boys born before the first girl is born
in hospitals 1, 2 and 3 respectively. If the observations are $X_{1}=0,X_{2}=3$
and $X_{3}=2$, find the most powerful test to test the null hypothesis
that a girl and a boy are equally likely to be born against the alternative
that a girl is less likely to be born than a boy.\hl{TODO}
\end{example}

\begin{example}
\label{exa:isi2005samplepsb10}Suppose that in 10 tosses of a coin
we get 7 heads and 3 tails. Find a test at level $\alpha=0.05$ to
test that the coin is fair against the alternative that the coin is
more likely to show up heads. Find the power function of this test.\hl{TODO}
\end{example}

\begin{example}
\label{exa:isi2006samplepsb12}Consider a possibly unbalanced coin
with probability of heads in each toss being $p$, where$p$ is unknown.
Let $X$ be the number of tails before the first head occurs. Find
the uniformly most powerful test of level $\alpha$ for testing $H_{0}:p=\frac{1}{6}$
against $H_{1}:p>\frac{1}{6}$.\hl{TODO}
\end{example}

\begin{example}
\label{exa:isi2007samplepsb11}Let $X_{1},X_{2},\ldots,X_{n}$ be
i.i.d. random variables from the exponential distribution with mean
$\theta>0$. Find the most powerful test for testing $H_{0}:\theta=2$
against $H_{1}:\theta=1$. Find the power of the test.\hl{TODO}
\end{example}

\begin{example}
\label{exa:isi2008samplepsb10}Consider a population with three kinds
of individuals labelled 1,2 and 3. Suppose the proportion of individuals
of the three types are given by $f(k,\theta),k=1,2,3$ where $0<\theta<1$
and 
\[
f(k,\theta)=\begin{cases}
\theta^{2} & \text{ if }k=1\\
2\theta(1-\theta) & \text{ if }k=2\\
(1-\theta)^{2} & \text{ if }k=3
\end{cases}
\]

Let $X_{1},X_{2},\ldots,X_{n}$ be a random sample from this population.
Find the most powerful test for testing $H_{0}:\theta=\theta_{0}$
versus $H_{1}:\theta=\theta_{1}\left(\theta_{0}<\theta_{1}<1\right)$
\hl{TODO}
\end{example}

\begin{example}
	\label{exa:isi2009samplepsb8}
	Let $X_1, \ldots, X_n$ be i.i.d. observations from the density
	$$
	f(x)=\frac{1}{\mu} \exp (-x / \mu), \quad x>0,
	$$
	where $\mu>0$ is an unknown parameter.
	Consider the problem of testing the hypothesis $\mathrm{H}_0: \mu \leq \mu_0$ against
	$$
	\mathrm{H}_1: \mu>\mu_0 .
	$$
	(a) Show that the test with critical region $\left[\bar{X} \geq \mu_0 \chi_{2 n, 1-\alpha}^2 / 2 n\right]$, where $\chi_{2 n, 1-\alpha}^2$ is the $(1-\alpha)$-th quantile of the $\chi_{2 n}^2$ distribution, has size $\alpha$.
	(b) Give an expression of the power in terms of the c.d.f. of the $\chi_{2 n}^2$ distribution.
	\hl{TODO}
\end{example}


